Parsing Postfix log files

### 2009-03-05 doc/logparser-architecture.tex
Corrected the formulae in the overlapping rules section, and corrected the formatting.

Overlapping rules can be detected by visual inspection, or a program may be
written to analyse the rule's regexes.  Traditional regexes are equivalent
in computational power to \acronym{FA} and can be converted to
\acronym{FA}, so regex overlap can be detected by finding a non-empty
intersection of two \acronym{FA}\@.  Perl 5.10 regexes~\cite{perlre} are
more powerful than traditional regexes: it is possible to match correctly
balanced brackets nested to an arbitrary depth, e.g.\
\verb!/^[^<>]*(<(?:(?>[^<>]+)|(?1))*>)[^<>]*$/!  matches
\verb!z<123<pq<>rs>j<r>ml>s!.  Matching balanced brackets requires the
regex engine to maintain state on a stack, so Perl 5.10 regexes are
equivalent in computational power to \acronym{PDA}; detecting overlap may
require calculating the intersection of two \acronym{PDA} instead of two
\acronym{FA}.  \acronym{PDA} intersection is not closed, i.e.\ the result
cannot always be implemented using a \acronym{PDA}, so intersection may be
intractable in some cases, e.g.:
$a^{*}b^{n}c^{n}~\cap~a^{n}b^{n}c^{*}~\rightarrow~a^{n}b^{n}c^{n}$.
Detecting overlap amongst $n$ regexes requires calculating $n(n-1)/2$
intersections, resulting in $O(n^2x)$ complexity, where $O(x)$ is the
complexity of calculating \acronym{FA} or \acronym{PDA} intersection.  This
is certainly not a task to be performed every time a parser runs: naive
detection of overlap amongst \parsernames{} \numberOFrules{} rules would
require calculating \numberOFruleINTERSECTIONS{} intersections.

@@

XXX CARL SUGGESTS COMPARING THIS TO THE POWERSET\@; IS THAT REASONABLE\@?

\noindent{}The first and second conditions do not overlap, but the third
condition overlaps with both the first and second conditions.  When rules
are divided into sets based on how their conditions overlap, the complexity
of detecting overlap amongst $n$ rules is:

XXX IN THE FORMULA BELOW $s \in{} overlaps$ SHOULD BE UNDER $\sum$

$$O(n^{2}y) + O\left(\sum{|s|^{2}x \forall{} s \in{} overlaps}\right)$$

where:

@@

            $x$ & = & cost of checking for overlap between two regexes    \\
            $y$ & = & cost of checking for overlap between two conditions \\
     $overlaps$ & = & set of sets of rules with overlapping conditions    \\

@@

For this approach to be more efficient than the naive approach $y$ must be
significantly lower than $x$.  If $y$ is higher than $x$ then the opposite
approach should be taken: check for regex overlap first, then check
for condition overlap only between rules that have overlapping regexes.

>>>>

Overlapping rules can be detected by visual inspection, or a program may be
written to analyse the rule's regexes.  Traditional regexes are equivalent
in computational power to \acronym{FA} and can be converted to
\acronym{FA}, so regex overlap can be detected by finding a non-empty
intersection of two \acronym{FA}\@.  Perl 5.10 regexes~\cite{perlre} are
more powerful than traditional regexes: it is possible to match correctly
balanced brackets nested to an arbitrary depth, e.g.\
\verb!/^[^<>]*(<(?:(?>[^<>]+)|(?1))*>)[^<>]*$/!  matches
\verb!z<123<pq<>rs>j<r>ml>s!.  Matching balanced brackets requires the
regex engine to maintain state on a stack, so Perl 5.10 regexes are
equivalent in computational power to \acronym{PDA}; detecting overlap may
require calculating the intersection of two \acronym{PDA} instead of two
\acronym{FA}.  \acronym{PDA} intersection is not closed, i.e.\ the result
cannot always be implemented using a \acronym{PDA}, so intersection may be
intractable in some cases, e.g.:
$a^{*}b^{n}c^{n}~\cap~a^{n}b^{n}c^{*}~\rightarrow~a^{n}b^{n}c^{n}$.
Detecting overlap amongst $n$ regexes requires calculating
$\frac{n\left(n-1\right)}{2}$ intersections, resulting in
$O\left(n^{2}x\right)$ complexity, where $x$ is the cost of calculating
\acronym{FA} or \acronym{PDA} intersection.  This is certainly not a task
to be performed every time a parser runs: naive detection of overlap
amongst \parsernames{} \numberOFrules{} rules would require calculating
\numberOFruleINTERSECTIONS{} intersections.

@@

\noindent{}The first and second conditions do not overlap, but the third
condition overlaps with both the first and second conditions.  When rules
are paired based on how their conditions overlap, the complexity of
detecting overlap amongst $n$ rules is $O\left(n^{2}y+|o|x\right)$, where:

@@

    $x$ & = & cost of checking for overlap between two regexes    \\
    $y$ & = & cost of checking for overlap between two conditions \\
    $o$ & = & set of pairs of rules with overlapping conditions   \\

@@

For this approach to be more efficient than the naive approach $y$ must be
significantly lower than $x$.  If $y$ is higher than $x$ then the opposite
approach should be taken: check for regex overlap first, then check
for condition overlap only between pairs of rules with overlapping regexes.
